---
title: "Medallists and medals of the 1896-2008 Summer Olympics"
author: "Radomska Małgorzata, Rapa Adrianna"
date: "2023-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
## Preparation of data for analysis

setwd("C:/Users/gosia/Desktop/Studia/Semestr II/Programowanie w R - projekt")

library(openxlsx)

medals <- read.xlsx("Summer_Olympic_medallists_1896-2008.xlsx")

unique_rows <- !duplicated(medals$Athlete)
medallists <- medals[unique_rows, ]

countries <- read.xlsx("Summer_Olympic_medallists_1896-2008.xlsx", 4)
head(countries1 <- countries[,1:2])

Int.Olympic.Committee.code <- c("ANZ", "BOH", "BWI","EUA", "EUN", 
                                "FRG", "GDR", "IOP", "ROU", "RU1", 
                                "SRB", "TCH", "URS", "YUG", "ZZX")
country_names <- c("Australasia", "Bohemia", "British West Indies", 
             "United Team of Germany","Unified Team", "West Germany", 
             "East Germany", "Independent Olympics Participants",
              "Romania", "Russian Empire", "Serbia", "Czechoslovakia",
             "Soviet Union","Yugoslavia", "Mixed team")

country_codes <- data.frame(country_names, Int.Olympic.Committee.code)
names(country_codes)[1] <- "Country"

countries2 <- rbind(countries1, country_codes)

for (i in 1:length(medals$NOC)){
  for (j in 1:length(countries2$Country)){
    if (medals$NOC[i]==countries2$Int.Olympic.Committee.code[j])
    {
      medals$NOC[i] <- countries2$Country[j]
    }
  }
}
```

### Number of medals from the countries from 1986 to 2008

```{r}
country <- table(medals$NOC)
country1 <- as.data.frame(country)

index <- which(country1$Freq >= 1000)
country2 <- country1[index, ]

library(ggplot2)

colour <- rep(c("dodgerblue2", "darkgoldenrod1", 
               "black", "springgreen3", "brown2"), length=nrow(country2))

country2$Var1 <- as.character(country2$Var1)

ggplot(country2, aes(x = Var1, y = Freq)) +
  geom_segment(aes(x = Var1, xend = Var1, y = 0, yend = Freq), color = colour) +
  geom_point(aes(color = Var1), size = 4) +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) +
  geom_text(aes(label = Freq, group = Var1), hjust = -0.75, vjust = 0.5) +
  labs(x = "Country", y = "Number of medals", 
       title = "Number of medals from the countries from 1986 to 2008") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = colour)
```


### The number of medallists whose names begin on consecutive letters of the alphabet

```{r}
letter <- c()

for (i in 1:nrow(medallists)){
  letter <- c(letter, substr(medallists$Athlete[i], 1, 1))
}

medallists_copy <- medallists

medallists_copy <- cbind(medallists_copy, letter)

gold <- medallists_copy[medallists_copy$Medal=="Gold", ]
silver <- medallists_copy[medallists_copy$Medal=="Silver", ]
bronze <- medallists_copy[medallists_copy$Medal=="Bronze", ]

gold_letter <- table(gold$letter)
gold_letter1 <- as.data.frame(gold_letter)
colnames(gold_letter1) <- c("Letter", "Gold")

silver_letter <- table(silver$letter)
silver_letter1 <- as.data.frame(silver_letter)
colnames(silver_letter1) <- c("Letter", "Silver")

bronze_letter <- table(bronze$letter)
bronze_letter1 <- as.data.frame(bronze_letter)
colnames(bronze_letter1) <- c("Letter", "Bronze")

all_colours_letter <- cbind(gold_letter1, silver_letter1[,2], bronze_letter1[,2])
colnames(all_colours_letter) <- c("Letter", "Gold", "Silver", "Bronze")

suppressMessages(suppressWarnings(library(tidyverse)))

letter_long <- all_colours_letter %>% gather(Medal, Number, -Letter)

ggplot(letter_long, aes(fill=Medal, y=Number, x=Letter)) + 
  geom_bar(position="dodge", stat="identity", width = 0.4) +
  scale_fill_manual(values = c("Gold" = "gold2", 
                               "Silver" = "grey", 
                               "Bronze" = "chocolate3")) +
  labs(x = "Letter", y = "Number of medals", 
       title = "The number of medallists whose names begin on consecutive letters of the alphabet") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Number of Polish medallists in consecutive years

```{r}
pol <- medals[medals$NOC=="Poland", ]

pol_gold <- pol[pol$Medal=="Gold", ]
pol_silver <- pol[pol$Medal=="Silver", ]
pol_bronze <- pol[pol$Medal=="Bronze", ]

pol_g <- table(pol_gold$Edition)
pol_g1 <- as.data.frame(pol_g)
colnames(pol_g1) <- c("Year", "Gold")

pol_s <- table(pol_silver$Edition)
pol_s1 <- as.data.frame(pol_s)
colnames(pol_s1) <- c("Year", "Silver")

pol_b <- table(pol_bronze$Edition)
pol_b1 <- as.data.frame(pol_b)
colnames(pol_b1) <- c("Year", "Bronze")

pol_medals <- merge(pol_b1, pol_s1, by="Year", all.x = TRUE, all.y = TRUE)
pol_medals1 <- merge(pol_medals, pol_g1, by="Year", all.x = TRUE, all.y = TRUE)

pol_medals1$Year <- as.numeric(as.character(pol_medals1$Year))

pol_medals2 <- pol_medals1[(order(pol_medals1$Year)), ]

pol_medals2$Year <- as.character(pol_medals2$Year)

pol_medals2[is.na(pol_medals2)] <- 0

pol_medals_long <- pol_medals2 %>% gather(Medal, Number, -Year)

ggplot(pol_medals_long, aes(fill=Medal, y=Number, x=Year)) + 
  geom_bar(position="stack", stat="identity", width = 0.5) +
  geom_text(aes(label=ifelse(Number != 0, Number, "")), 
            position=position_stack(vjust = 0.5), color="black", size=3) +
  scale_fill_manual(values = c("Bronze" = "chocolate3", 
                               "Silver" = "grey", 
                               "Gold" = "gold2")) +
  labs(x = "Year", y = "Number of medallists", 
       title = "Number of Polish medallists in consecutive years")+
  theme(plot.title = element_text(hjust = 0.5))
```

### Percentage summary of Polish medallists in each discipline

```{r}
pol_gold_discipline <- table(pol_gold$Discipline)
pol_g_discipline <- as.data.frame(pol_gold_discipline)
colnames(pol_g_discipline) <- c("Discipline", "Gold")

pol_silver_discipline <- table(pol_silver$Discipline)
pol_s_discipline <- as.data.frame(pol_silver_discipline)
colnames(pol_s_discipline) <- c("Discipline", "Silver")

pol_bronze_discipline <- table(pol_bronze$Discipline)
pol_b_discipline <- as.data.frame(pol_bronze_discipline)
colnames(pol_b_discipline) <- c("Discipline", "Bronze")

pol_discipline <- merge(pol_b_discipline, pol_s_discipline, by="Discipline", all.x = TRUE, all.y = TRUE)
pol_discipline1 <- merge(pol_discipline, pol_g_discipline, by="Discipline", all.x = TRUE, all.y = TRUE)

pol_discipline1[is.na(pol_discipline1)] <- 0

pol_discipline_long <- pol_discipline1 %>% gather(Medal, Number, -Discipline)

ggplot(pol_discipline_long, aes(fill=Medal, y=Number, x=Discipline)) + 
  geom_bar(position="fill", stat="identity", width = 0.5) +
  scale_fill_manual(values = c("Bronze" = "chocolate3", 
                               "Silver" = "grey", 
                               "Gold" = "gold2")) +
  labs(x = "Discipline", y = "Medallists", 
       title = "Percentage summary of Polish medallists in each discipline")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

```

### Number of medallists at consecutive Olympic Games by gender

```{r}

suppressMessages(suppressWarnings(library(data.table)))

medallists <- as.data.table(medallists)

medallists_gender_count <- medallists[, .N, by = .(Edition, Gender)]

medallists_gender <- dcast(medallists_gender_count, Edition ~ Gender, value.var = "N")
medallists_gender <- medallists_gender[, .(Edition, Women, Men)]

medallists_gender[is.na(medallists_gender)] <- 0

ggplot(medallists_gender_count, aes(fill = Gender, y = N, x = as.factor(Edition))) +
  geom_bar(position = "dodge", stat = "identity", width = 0.8) +
  geom_text(aes(label = N), position = position_dodge(width = 0.8), vjust = -0.5, angle = 90, hjust = 0.5, size = 3) +
  scale_fill_manual(values = c("Women" = "lightpink", "Men" = "lightblue3")) +
  labs(x = "Year", y = "Number", 
       title = "Number of medallists at consecutive Olympic Games by gender") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(medallists_gender_count$N), by = 100))

```

### The number of participants who have won a medal in a specific edition in different disciplines

```{r}

selected_columns <- medals %>%
  select(Edition, Discipline, Athlete, Gender, Event, Medal)

few_medals_won <- data.frame(Edition = medals$Edition,
                                             Discipline = medals$Discipline,
                                             Athlete = medals$Athlete,
                                             Gender = medals$Gender,
                                             Event = medals$Event,
                                             Medal = medals$Medal)

repeated <- few_medals_won[duplicated(few_medals_won[, c("Athlete", "Edition", "Medal")]) |
            duplicated(few_medals_won[, c("Athlete", "Edition", "Medal")], fromLast = TRUE), ]

repeated_sorted <- repeated %>% 
  arrange(Edition, Athlete)

situation_count <- repeated_sorted %>% 
  group_by(Edition, Athlete) %>% 
  count()

participant_count <- situation_count %>%
  group_by(Edition) %>%
  summarize(participants_count = n_distinct(Athlete))

participant_count$Edition <- factor(participant_count$Edition)

ggplot(participant_count, aes(x = Edition, y = participants_count, fill = Edition)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d() +
  labs(x = "Edition", y = "Number of participants", fill = "Edition", 
  title = "The number of participants who have won a medal in a specific edition 
           in different disciplines") +
  theme_minimal() +
  theme(legend.position = "right", legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(breaks = seq(0, max(participant_count$participants_count), by = 10),
                     minor_breaks = seq(0, max(participant_count$participants_count), by = 10),
                     limits = c(0, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Number of situations in which several of the same medals were awarded for identical results for each edition

```{r}

selected_columns <- medals %>%
  select(Edition, Discipline, Event, Athlete, Medal)

repeated_medals <- selected_columns %>%
  group_by(Edition, Discipline, Event, Medal) %>%
  summarise(Number = n()) %>%
  filter(Number >= 2)

event_count <- repeated_medals %>%
  group_by(Edition) %>%
  summarise(event_count = n_distinct(Event))

repeated_color_medals <- repeated_medals

repeated_color_medals$Kolor <- ifelse(repeated_color_medals$Medal == "Gold", "Gold",
                                      ifelse(repeated_color_medals$Medal == "Silver", "Silver",
                                             ifelse(repeated_color_medals$Medal == "Bronze", "Bronze", "Other")))

event_count <- repeated_color_medals %>%
  group_by(Edition) %>%
  summarise(Number_of_situations = n_distinct(paste(Discipline, Event, Kolor)))

ggplot(event_count, aes(x = Edition, y = Number_of_situations)) +
  geom_point(size = 3, color = "darkgreen") +
  scale_x_continuous(breaks = event_count$Edition, labels = event_count$Edition) +
  labs(title = "Number of situations in which several of the same medals were awarded for identical results for each edition",
       x = "Edition",
       y = "Number of situation") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Number of situations in which several of the same medals were awarded for identical results for each edition")

```

### Medallists who have won several medals of the same colour in different editions

```{r}
selected_columns2 <- medals %>%
  select(Edition, Athlete, Discipline, Event, Medal)

selected_columns3 <- selected_columns2 %>%
  arrange(Athlete, Discipline, Edition, Medal)

grouped_data <- selected_columns3 %>%
  group_by(Athlete, Event)

dup_event_participants <- grouped_data %>%
  filter(n_distinct(Edition) > 1)

grouped_data <- dup_event_participants %>%
  group_by(Athlete, Event, Medal)

filtered_participants <- grouped_data %>%
  filter(n_distinct(Edition) > 1)

grouped_data <- filtered_participants %>%
  group_by(Athlete, Event, Medal)

counted_editions <- grouped_data %>%
  summarise(Count = n_distinct(Edition))

sorted_data <- counted_editions %>%
  arrange(desc(Count))

top_participant <- sorted_data[1:5, ]

winning_records <- filtered_participants %>%
  filter(Athlete %in% top_participant$Athlete[1:5])

selected_columns <- winning_records %>%
  select(Athlete, Medal, Event)

medal_counts <- winning_records %>%
  group_by(Athlete, Event, Medal) %>%
  summarise(MedalCount = n())

medal_colors <- c("Gold" = "gold2", "Silver" = "grey", "Bronze" = "chocolate3")

value_plot <- ggplot(medal_counts, aes(x = reorder(paste(Athlete, Event), -MedalCount), y = MedalCount, fill = Medal)) +
  geom_col(width = 0.7, position = position_dodge()) +
  geom_text(aes(label = MedalCount), vjust = -0.5, size = 4, color = "black", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = medal_colors) +
  labs(x = "Athlete / Cztegory", y = "Number of medals", fill = "Medal",
       title = "Medallists who have won several medals of the same colour in different editions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))

print(value_plot)
```

### Countries with the highest number of female participants in the respective disciplines

```{r}
women <- medallists %>%
  select(Edition, Discipline, Event, NOC, Gender)

count_women <- medallists %>%
  filter(Gender == "Women") %>%
  group_by(Discipline,NOC, Event) %>%
  summarize(Women_Count = n())

count_women_sorted <- count_women %>%
  arrange(Discipline, Event, NOC, Women_Count)

top_countries <- count_women %>%
  arrange(Discipline, Event, desc(Women_Count)) %>%
  group_by(Discipline, Event) %>%
  top_n(1, Women_Count)

top_countries2 <- count_women %>%
  arrange(Discipline, desc(Women_Count)) %>%
  group_by(Discipline) %>%
  slice_max(Women_Count)

ggplot(top_countries2, aes(x = Women_Count, y = Discipline, fill = NOC)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Countries with the highest number of female participants in the respective disciplines",
       x = "Number of women",
       y = "Discipline") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.title = element_text(hjust = 0.5))

```